---
title: "Progress Report"
author: "Varun Yeduru"
date: "04/22/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load necessary libraries
library(rvest)
library(dplyr)
library(Lahman)

# Specify the URL for the 2023 Hall of Fame data
url <- "https://www.baseball-reference.com/awards/hof_2023.shtml"

# Retrieve and parse the HTML from the URL
web_data <- read_html(url)

# Isolate the table with Hall of Fame voting results
voting_data_table <- web_data %>% html_table(fill = TRUE)

# Extract the first table as it contains our data
hof_voting_data <- voting_data_table[[1]]

# Display the structure of the data we extracted
str(hof_voting_data)

# Use Lahman for comparison purposes
head(HallOfFame, 5)



```
```{r}
# Load essential packages
library(rvest)
library(dplyr)
library(Lahman)

# Define the URL for the 2023 Hall of Fame data
url <- "https://www.baseball-reference.com/awards/hof_2023.shtml"

# Fetch and parse the HTML content from the webpage
web_content <- read_html(url)

# Extract the voting data table
data_table <- web_content %>% html_table(fill = TRUE)

# Select the table presumed to contain Hall of Fame data
voting_results_2023 <- data_table[[1]]

# Standardize column names to align with the Lahman database's structure
colnames(voting_results_2023) <- c("playerID", "yearID", "votedBy", "ballots", "needed", "votes", "inducted", "category", "additionalNotes")

# Display unique values to identify non-numeric entries
unique(voting_results_2023$yearID)
unique(voting_results_2023$ballots)
unique(voting_results_2023$needed)
unique(voting_results_2023$votes)

# Modify data types for proper analysis
voting_results_2023$yearID <- as.integer(voting_results_2023$yearID)
voting_results_2023$ballots <- as.integer(voting_results_2023$ballots)
voting_results_2023$needed <- as.integer(voting_results_2023$needed)
voting_results_2023$votes <- as.integer(voting_results_2023$votes)
voting_results_2023$inducted <- ifelse(voting_results_2023$inducted == "Y", "Y", "N")

# Combine the newly cleaned data with the existing HallOfFame data
updated_hall_of_fame <- bind_rows(HallOfFame, voting_results_2023)

# Save the enhanced dataset
save(updated_hall_of_fame, file = "Updated_HallOfFame.rda", compress = "xz")








```
```{r}
# Visualize the data with ggplot2
library(ggplot2)
updated_hall_of_fame %>% 
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2023)) +
  labs(title = "Trends in Hall of Fame Inductions", x = "Year of Induction", y = "Number of Inductees")


```

displays the number of inductees into a Hall of Fame over the years, spanning from the earlier 20th century through approximately 2023. The bars are divided by color to indicate whether individuals were inducted (blue) or not inducted (red).

From the chart, we can observe this:

There are notable fluctuations in the number of inductees year over year.
In some years, there are spikes in the number of inductees, while other years see relatively few or no inductees.
The red bars (non-inductees) are more prevalent throughout the years, suggesting that in any given year, the number of nominees who do not get inducted surpasses those who do.

```{r}
# Export the updated Hall of Fame data to a CSV file
write.csv(updated_hall_of_fame, "Updated_HallOfFame2023.csv", row.names = FALSE)

```

